### 2016-01-21 Manuscript/BiologyLetters/response.tex

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{arabic}

%Pagination style and stuff
%\linespread{2}

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: MPE-15-254}\\
\bigskip
Dear Editor,\\ % put editor's name
\bigskip
We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper. We have taken all of their comments on board, and respond to their points below. For improving clarity in this document, we sorted the reviewers comments as follows:
\begin{enumerate}
\item \textbf{Reviewer 1:} containing reviewer 1's major comments
\item \textbf{Reviewer 1 specific comments:} containing reviewer 1's specific comments
\item \textbf{Reviewer 2:} containing reviewer 2's comments
\end{enumerate}
Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.

\section{Reviewer 1:}

\begin{enumerate}
% -----------------
% Comment 1
% -----------------
\item{\textcolor{blue}{Coment 1}}

\end{enumerate}

We hope we have responded to all these comments appropriately. Please let us know if you require any further information,\\
\bigskip

Thomas Guillerme (guillert@tcd.ie)\\
Natalie Cooper

\end{document}

### 2016-01-21 Manuscript/BiologyLetters/response.tex

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: MPE-15-254}\\
\bigskip
Dear Surayya Johar,\\
\bigskip
We are very grateful to the four referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper. We have taken all of their comments on board, and respond to their points below. For improving clarity in this document, we dealt with each comment in the order they appeared in your ``Decision on Manuscript'' email and we copied the reviewers comments in blue.

%%%%%%%%%%%%%%%
% Reviewer 1
%%%%%%%%%%%%%%%

\section{Reviewer 1:}
\begin{enumerate}
\item{\textcolor{blue}{In ESM 1, p. 4, the authors explain their methods of collecting morphological data from the literature. Why select only the first 10 results per search term? How do you know that relevant papers wouldn’t come up later in the search? I need evidence to support this cutoff; as it stands, I’m concerned that you’re missing some relevant material.}}

\item{\textcolor{blue}{Similarly, I find the 2010 date cutoff troubling, because I’m not sure that the argument that these recent studies contain a fraction of the previously published characters and OTUs allows you to make a comprehensive statement about the availability of character data. What about the fraction that isn’t included? This is a question that needs to be better addressed before you will convince readers that this is really all the data that is out there.}}
\end{enumerate}

\subsection{Reviewer 1 specific comments:}
\begin{enumerate}
\item{\textcolor{blue}{P.2, line 43: ``chunks'' is not an appropriate term to use in a scientific paper. Use a more formal term.}}
\item{\textcolor{blue}{P.2, L53 \& L55: ``branch'' is the wrong verb. You want to say something like ``fossils will not be placed accurately within the phylogeny'' }}
\item{\textcolor{blue}{P3, L78: ``since'' should be ``because.'' Since should be used for time, not causation.}}
\item{\textcolor{blue}{P3, L85: ``not-applicable'' should not be hyphenated.}}
\item{\textcolor{blue}{P3, L120: How much of a difference does the 25\% or less coverage make to the capacity of Total Evidence phylogenetic analyses to resolve relationships? Does your simulation study have a way of quantifying or describing the probability of getting the ``right'' answer with this little data? It would also help if you could show in your data table what the clades were that had better than 75\% coverage. It’s not immediately evident from reading Table 1.}}
\item{\textcolor{blue}{P5, L137: ``taxa'' should be ``OTUs.''}}
\item{\textcolor{blue}{P5, L165 \& 166: This sentence also misuses ``branch'' as on P.2. Reword for correct verb usage.}}
\item{\textcolor{blue}{ESM 1, p. 2: ``If some where present'' should be ``if some were present.''}}
\item{\textcolor{blue}{ESM 1, p. 3: ``one of these repository'' should be ``one of these repositories.''}}
\item{\textcolor{blue}{ESM 1, p. 4: ``that were to irrelevant'' should be ``that were too irrelevant.''}}
\item{\textcolor{blue}{ESM 1, p. 4: ``of the recent published matrix'' should be ``of the recently published matrices.''}}
\item{\textcolor{blue}{ESM 1, p. 4: First sentence of the paragraph in the middle of the page is a sentence fragment, and needs to state where the matrices are available.}}
\item{\textcolor{blue}{ESM 1, p. 4: ``since the entries in the…'' should be ``because the entries in the…'' Since refers to time, not causation.}}
\item{\textcolor{blue}{ESM 1, p. 5: axis title should be ``Google Scholar matches,'' not ``Google Scholars matches.'' Figure caption should capitalize ``Google Scholar.'' Y-axis and X-axis should be hyphenated.}}
\item{\textcolor{blue}{ESM 1, p. 6: ``wrong binomial names format'' should be ``incorrectly-formatted binomial names,'' and ``into the correct ones'' should be deleted. Fixing is assumed to turn incorrect things into correct ones. Also, the sentence starting the paragraph under ``Selecting the living OTUs'' repeats the exact same thing as this sentence.}}
\item{\textcolor{blue}{ESM 2 seems to be almost completely redundant with Table 1. Is there some more efficient way of dealing with this information?}}
\end{enumerate}

%%%%%%%%%%%%%%%
% Reviewer 2 (Graham Slater)
%%%%%%%%%%%%%%%

\section{Reviewer 2:}
\begin{enumerate}
\item{\textcolor{blue}{The authors state on page 3 lines 82-90 that they avoided matrices with ``few characters'' due to lower probability of overlapping characters with other matrices and so only selected matrices containing >100 characters. [...] Is 100 really the minimum number you want to go with at all phylogenetic scales?}}
\item{\textcolor{blue}{Going through each clade in the supplement seems beyond the scope of reviewing duties, but I am concerned that if this one matrix can be missed others may have been missed too. I think a reexamination of the literature may be warranted using taxon specific search terms.}}
\item{\textcolor{blue}{One interesting aside that the authors didn’t consider, despite the data being there, is how phylogenetic ``effort'' relates to diversity of the order under study.}}

We are really grateful for Graham Slater's %or do we need to keep it formal and say just "reviewer 2's"
comment and added this interesting analysis in the ESM @@@% get number.
% that mammalian morphological phylogeneticists have focused their energies on resolving the higher level relationships among mammalian lineages.
\end{enumerate}

%%%%%%%%%%%%%%%
% Reviewer 3 (Peter Wagner)
%%%%%%%%%%%%%%%

\section{Reviewer 3:}
\begin{enumerate}
\item{\textcolor{blue}{My biggest criticism is a semantic one: Morphological data $\neq$ cladistic data.}}
\item{\textcolor{blue}{The real issue here is the availability of phylogenetic character coding for anatomical data.
[...] Thus, the title really should be ``Assessment of available anatomical characters for phylogenetic analysis among living mammals'' or something like that.}}
\item{\textcolor{blue}{what we really need is to focus on the tooth and skull characters that typify the fossil record.
[...] Still, several papers estimate that the sampling of mammal species (based on cranial characters include teeth) actually is pretty darned good [...].
Thus, we might be able to richly populate FBD and tip-dating analyses with a LOT of extinct species.}}
\item{\textcolor{blue}{Finally, I would add: who the heck submits a paper for a special volume 6 weeks early?!?!? You make the rest of us look bad!}}
This paper was actually ready to submit from the 28th of July (\url{http://biorxiv.org/content/early/2015/07/28/022970}) which partially explain the swift submission.
We do sincerely apologize for making the rest of the authors submitting to the issue look bad.

@@

\end{enumerate}

\subsection{Reviewer 3 specific comments:}
\begin{enumerate}
\item{\textcolor{blue}{Lines 25-26: ``We suggest that increased morphological data collection efforts for living taxa are needed to produce accurate Total Evidence phylogenies.''
In a way, this is more than ``total evidence.=''.
The initial total evidence paradigm was simply to use every character you could (molecular, morphological, even behavioral).
We now are talking about tip-dating and fossilized birth-death (FBD) approaches (e.g., Heath et al. 2014 PNAS 111:E2957) that want to use all of the available data.
After all, one can do a total evidence analysis with just node-dating techniques rather than with tip-dating, and one can do it without worrying about branch length priors.
However: if we want to use all of the data to calibrate branch lengths for all taxa (fossil and extant), then we need ``total evidence'' to do tip-dating and FBD properly).
As tip-dating is the main theme of this issue, that should be linked to the total evidence explicitly here and elsewhere in the manuscript.}}

\item{\textcolor{blue}{Lines 70-71: ``We downloaded all cladistic matrices containing any living and/or fossil mammal  taxa from three major public databases:''
As a side note and for future reference for the authors, a large number of phylogenetic data sets for mammals (and other metazoans) also are available at the Paleobiology Database.
(These were formerly at my website at the Field Museum, but that site obviously disappeared when I moved to the Smithsonian.)
Because of the nature of the projects for which I was using these datasets, most of them are species-level, but some are genus-level, too.}}

\item{\textcolor{blue}{Lines 164-166: ``For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure 1B).'' 
This seems to be an awkward way to phrase this.
Moreover, this seems like an extension of the long-branch problem.
Basically, the issue really is that we do not have coded skeletal characters for extant herpestids.
Thus, it is not possible for coded skeletal characters from fossil herpestids to connect those fossil herpestids to the extant herpestids.
This is the opposite of Felsenstein’s long branch: insofar as skeletal characters are concerned, extant herpestids have zero-length branches.
That in turn means that the fossil herpestids are going to be ``attracted'' to non-zero length branches elsewhere, with long branches in related groups such as canids being particularly good candidates because there is a higher chance of homoplasies between canids and herpesitds on the longer branches.
So, it will not be random: they probably will latch themselves onto the longest reasonably branch among the closest relatives for which anatomical data are present.}}

%%%%%%%%%%%%%%%
% Reviewer 4 (Graeme Lloyd)
%%%%%%%%%%%%%%%

\section{Reviewer 4:}
\begin{enumerate}
\item{\textcolor{blue}{One point I would add of my own is that the orders used are obviously very variable in size (species number), so what about total coverage?
This could surely be synthesised from the tables and would be a nice simple way to compare the with and without threshold results.
Similarly, I think Reviewer 3’s [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
 graphs are a nice potential addition to the ESM.}}
\end{enumerate}

\subsection{Reviewer 4 specific comments:}
\begin{enumerate}
\item{\textcolor{blue}{Lines 72-74 - Web addresses should be consistent in dropping of ``http://www.''.}}
\item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
\item{\textcolor{blue}{Line 104 - It’s not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
\item{\textcolor{blue}{Line 117 - It’s not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level. Are you using the node subtending the higher taxon or something else?}}
\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I’m not sure if this is strictly true. It’s more that shorter OTU names are commonly used. After all at least one actual specimen has to be coded and in my experience if you read the paper usually only a single species is used. Palaeontologists are just bad at using proper binomials in their matrices (which is an issue, but a different one). In addition ref 20 is actually all species-level OTUs! (I didn’t check ref 21.)}}
\item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}
\item{\textcolor{blue}{Line 245 - The caption could be more informative regarding the random and clustered distributions that these figures represent.}}
\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}
\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}
\item{\textcolor{blue}{ESM 1 - ``designated as ``fossil'' all OTUs that were present in the Paleobiology database'' The database contains extant taxa too, although these should be marked as such. I guess this doesn’t matter given the flow chart in Figure 2, but is something the authors should be aware of. (Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}
\item{\textcolor{blue}{ESM 1 Figure 2 - How many OTUs were ignored? And based on the (assumed) cursory inspection of the authors what are they? Higher taxa? Specimens? Or just species you can’t reconcile with your three data sets?}}
\item{\textcolor{blue}{ESM 2 - ``The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric'' Abbreviations should be introduced when first used, not later as they are here.}}
\end{enumerate}

@@

We hope we have responded to all these comments appropriately. Please let us know if you require any further information,\\
\bigskip

% Editorial office comments to authors:
% Please ensure that you include; 
% *An author contributions section which fulfils all *four* of our criteria given here https://royalsociety.org/journals/ethics-policies/openness/, including a sentence stating all authors agree to be held accountable for the content therein and approve the final version of the manuscript 
% *A funding statement which is the same as that given in ScholarOne
% *Please also confirm in your cover letter whether all the figures are your own or whether permission has been obtained for their use
% *Please upload your original figure files as eps, tiff or jpeg files, rather than PDFs
% *If you have any images that can be used to promote your article on social media (should this be accepted) please upload them as a supplementary file
% *Please confirm whether your manuscript should be available via open access, as we note that this may be a requirement of your funder.

Thomas Guillerme (guillert@tcd.ie)\\ % Changing email + adding imperial address?
Natalie Cooper

### 2016-01-21 Manuscript/BiologyLetters/Missing_mammals-draft.tex
response to minor comments reviewer 1

There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g. \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips. 
A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa.
Chunks of this data can be difficult, or impossible, to collect for every taxon in the analysis.
For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
Additionally, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g. in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded morphological data).
Unfortunately this missing data can lead to errors in phylogenetic inference.
Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between morphological data in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few morphological data \cite{GuillermeCooper}.
This is because (1) fossils cannot branch in the correct clade if it contains no morphological data for living taxa; and (2) fossils have a higher probability of branching within clades with more morphological data for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.

@@

%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
We used OTUs rather than species since entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.

Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not-applicable across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
Results of analyses with no threshold are available in the ESM. 
After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)

@@

@@

The number of living mammalian taxa with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data.
This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
Therefore, most palaeontological studies use the genus as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to branch in the Herpestidae, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).

@@

%Biology letters various stuff
\section{Ethics statement}
N/A
\section{Data accessibility statement}
All data and code are available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
\section{Authors' Contributions}
T.G. and N.C designed the study. T.G. analysed the data. T.G. and N.C. wrote the the manuscript.
\section{Competing Interests}
We have no competing interests.
\section{Acknowledgments}
We thank David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
\section{Funding statement}
This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

>>>>

There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix (e.g. \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a phylogeny with living and fossil taxa at the tips. 
A downside of this method is that it requires molecular data for living taxa and morphological data for both living and fossil taxa.
Parts%rev 1.2.1 (as in reviewer 1, part 2 (minor), comment 1)
 of this data can be difficult, or impossible, to collect for every taxon in the analysis.
For example, fossils rarely have molecular data and incomplete fossil preservation may restrict the amount of morphological data available.
Additionally, it has become less common to collect morphological characters for living taxa when molecular data is available (e.g. in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded morphological data).
Unfortunately this missing data can lead to errors in phylogenetic inference.
Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between morphological data in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few morphological data \cite{GuillermeCooper}.
This is because (1) fossils will not be placed accurately within the correct clade% rev 1.2.2
 if it contains no morphological data for living taxa; and (2) fossils have a higher probability of branching within clades with more morphological data for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.

@@

%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
We used OTUs rather than species because% rev 1.2.3
 entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.

Matrices with few characters are problematic when comparing available data among matrices because (1) they have less chance of having overlapping characters with other matrices \cite{wagner2000} and (2) they are more likely to contain specific characters that are not applicable% rev 1.2.4
 across large clades (e.g. ``antler ramifications'' is a character that is only applicable to Cervidae \cite{Brazeau2011}).
Therefore we selected only matrices containing \textgreater 100 characters for each OTU.
This threshold was chosen to correspond with the number of characters used in \cite{GuillermeCooper} and \cite{harrisonamong-character2014}.
Results of analyses with no threshold are available in the ESM. 
After removing matrices with \textless 100 characters, we retained 1074 unique living OTUs from 126 matrices. % 1601 unique living OTUs for 286 matrices (no threshold)

@@

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{example_coverage.pdf}
\caption{Dummy figure.}
\label{Figure_example_coverage}
\end{figure}

@@

The number of living mammalian OTUs% rev 1.2.6
 with no available cladistic data was surprisingly high at the species-level: only six out of 28 orders have a high coverage of taxa with available cladistic data.
This high coverage threshold of 75\% of taxa with available cladistic data represents the minimum amount of data required before missing data has a significant effect on the topology of Total Evidence trees \cite{GuillermeCooper}.
Beyond this threshold, there is considerable displacement of wildcard taxa and decreased clade conservation \cite{GuillermeCooper}.
Therefore we expect difficulties in placement of fossils at the species-level in most mammalian orders, but fewer issues at higher taxonomic levels.
This point is important from a practical point of view because of the slight discrepancy between neontological and palaeontological species concepts.
While neontological species are described using morphology, genes, distribution etc.; palaeontological species can be based only on morphological, spatial and temporal data (e.g. \cite{ni2013oldest}).
Therefore, most palaeontological studies use the genus as their smallest OTU (e.g. \cite{ni2013oldest,O'Leary08022013}), so data availability at the genus-level in living mammals should be our primary concern when building phylogenies of living and fossil taxa.

When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).

@@

%Biology letters various stuff
\section{Ethics statement}
N/A
\section{Data accessibility statement}
All data and code are available on GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
\section{Authors' Contributions}
T.G. and N.C designed the study. T.G. analysed the data. T.G. and N.C. wrote the the manuscript.
\section{Competing Interests}
We have no competing interests.
\section{Acknowledgments}
We thank David Bapst, Graeme Lloyd, Nick Matzke and April Wright.
Thanks to one anonymous reviewer, Graham Slater, Peter Wagner and Graeme Lloyd for their useful comments.
%That's two times Graeme though...
\section{Funding statement}
This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

### 2016-01-21 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
response to minor comments reviewer 1

Details about each public repository specific search option is listed below.
Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some where present in more than one matrix, they still only counted as one single OTU.

@@

\subsection{Google scholars (accessed 11th of June 2015)}
To make sure we didn't miss any extra matrix that wasn't available on one of these repository, we ran a Google Scholar search on the 5th of January. 
We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
We used the following key words:

@@

were \textit{order} was replaced by all the keywords listed above. For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
We selected only the 20 first results per search term to avoid downloading articles that were to irrelevant. Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_google_searches.pdf}
\caption{Google searches additional OTUs rarefaction curve. The x axis represent the number of google scholar matches (papers, books or abstracts) and the y axis represents the cumulative number of additional living OTUs per google scholar match.}
\label{Supp_figure_google_searches}
\end{figure}

The list of all the 286 download matrices is available on % PREMALINK TO THE MATRICES LIST (TRANSFORM TO CSV!)
The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
In this study, we refer to OTUs rather than species since the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e. the name of a collection item) to the family-level.
Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genera than species.

\subsection{Standardising the matrices}
We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
We then manually fixed the wrong bionomial names format (e.g. \textit{H. sapiens}) into the correct ones (e.g. \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
All the standardised matrices are available on % PREMALINK TO THE STANDARDISED MATRICES

>>>>

Details about each public repository specific search option is listed below.
Note that some matrices have been downloaded from more than one database but that it is not an issue since we are interested in the total number of unique living OTUs and that if some were present in more than one matrix, they still only counted as one single OTU.

@@

\subsection{Google scholars (accessed 11th of June 2015)}
To make sure we didn't miss any extra matrix that wasn't available on one of these repositories, we ran a Google Scholar search on the 5th of January. 
We downloaded the additional cladistic matrices from the 20 first search results matching with our selected keywords and with any of the 35 taxonomic levels (mammals Orders, Infraclasses and Class).
We used the following key words:

@@

were \textit{order} was replaced by all the keywords listed above. For each 33 keywords, we selected the 20 first papers to match the Google search published since 2010 resulting in 660 papers.
Among these papers, not all contained relevant data (discrete morphological characters AND mammalian data).
We selected only the 20 first results per search term to avoid downloading articles that were too irrelevant. Among the 660 papers, only 50 contained a total of 425 extra living OTUs (Figure ~\ref{Supp_figure_google_searches}).
Also we decided to select only the articles published since 2010 because nearly every one of the recent published matrix contains both a fraction of morphological characters and OTUs from previous studies.
For example in primates the character \textit{p7} coded first by \cite{ross1998phylogenetic} is reused with the same living species in \cite{seiffert2003fossil}, \cite{marivaux2005anthropoid}, \cite{seiffert2005basal}, \cite{bloch2007new}, \cite{bloch2007new}, \cite{kay2008anatomy}, \cite{silcox2008biogeographic}, \cite{seiffert2009convergent}, \cite{tabuce2009anthropoid}, \cite{boyer2010astragalar}, \cite{seiffert2010fossil}, \cite{marivaux2013djebelemur} and \cite{ni2013oldest}.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_google_searches.pdf}
\caption{Google searches additional OTUs rarefaction curve. The x-axis represent the number of Google Scholar matches (papers, books or abstracts) and the y-axis represents the cumulative number of additional living OTUs per google scholar match.}
\label{Supp_figure_google_searches}
\end{figure}

The list of all the 286 download matrices is available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
In this study, we refer to OTUs rather than species because the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e. the name of a collection item) to the family-level.
Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genera than species.

\subsection{Standardising the matrices}
We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
We then manually fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
All the standardised matrices are available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
